### 2014-03-05 src/x1x32_perline_example.tex

killed:
./AntiCheat/src/main/java/net/h31ix/anticheat/xray/XRayTracker.java

        return calculate(player, stats.getDiamond(), total) || calculate(player, stats.getGold(), total) || calculate(player, stats.getIron(), total) || calculate(player, stats.getLapis(), total) || calculate(player, stats.getRedstone(), total);

live:
./AntiCheat/src/main/java/net/h31ix/anticheat/util/PastebinReport.java
 out.write(("api_option=paste" + "&api_dev_key=" + URLEncoder.encode("c0616def494dcb5b7632304f8c52c0f1", "utf-8") + "&api_paste_code=" + URLEncoder.encode(report.toString(), "utf-8") + "&api_paste_private=" + URLEncoder.encode("1", "utf-8") + "&api_paste_name=" + URLEncoder.encode("", "utf-8") + "&api_paste_expire_date=" + URLEncoder.encode("1M", "utf-8") + "&api_paste_format=" + URLEncoder.encode("text", "utf-8") + "&api_user_key=" + URLEncoder.encode("", "utf-8")).getBytes());