### 2014-03-09 src/intro.Rnw
update

\section{Introduction}

% 1. Describe the problem
% 2. State your contributions
Mutation analysis is a method of evaluation of quality of test-suites. It involves producing
a family of mutants with simple differences from the original program, and evaluating the
effectiveness of test-suites against the these mutants~\cite{lipton1971fault, ammann2008introduction}.
Previous research~\cite{andrews2005mutation} suggests that mutations thus introduced behave
in a fashion similar to the real-faults with respect to the difficulty of detection.

One of the impediments to wider adoption of mutation analysis is its high computational cost.
The set of simple mutants of a moderate sized program can be very large, and renders
mutation analysis prohibitively time-consuming.

A major strain of research into cost-reduction of mutation analysis is to choose a smaller
representative set of mutants~\cite{offutt2001mutation,jia2011analysis}-- often called the
``do fewer'' approach. This approach can be generally divided into selective strategies
and sampling strategies.

Selective mutation strategies attempt to select a representative subset of operators based
on heuristics and statistical analysis, and applying this subset of operators to generate mutants
instead of applying the whole gamut of mutation operators~\cite{offutt1993anexperimental,untch2009onreduced}.

Sampling strategies seek to randomly select smaller representative mutants.
This was investigated first by Acree~\cite{acree1980mutation} and Budd~\cite{budd1980mutation},
who proposed using only $x$\% of the mutants produced. The studies by Wong and
Mathur found that random sampling with ratio as low as 10\% could provide
accurate results~\cite{jia2011analysis}.

%Mathur \cite{} reduced the number of mutagens in Mothra \cite{Mothra} by omitting two operators.
%Offutt et al.\cite{offutt1993anexperimental} further omitted upto six oprators from the
%mutagens and observed negligible loss of effectiveness in mutation analysis.
%Unch\cite{untch2009onreduced} and more recently Deng et al.\cite{deng2013empirical} suggest
%that using only statement deletion operator gives a close approximation of original mutation score.

%While some studies try to
%find the \emph{sufficient} subset of mutagens, like\cite{siami2008sufficient}, others
%have suggested that \emph{random} selection of mutants works as fine, e.g. \cite{zhang2010operator}.
%Some approaches combine the selective mutation selection and random selection, e.g. \cite{zhang2013ase}.

% {Amin: I am not a fan of this ~\cite{hussain08mutation} }
%Finally, clustering of mutants~\cite{hussain08mutation} based on similar test cases has also
%been used successfully for mutant reduction. These are explained more in depth in Section~\ref{sec:related}

A few recent works~\cite{zhang2010icse,zhang2013ase} have investigated the relative merits of
random sampling strategies and operator selection. It was found that while random sampling can
generally perform equal to or better than operator selection, a strategy that combined random
sampling with program elements paid superior dividends.

However, as pointed out by Zhang et al.~\cite{zhang2013ase}, the field has a serious lacuna in
large scale research, both in the size of the programs studied, and in the number and diversity
of programs, which reduces our confidence in the espoused results. This is true both for
selective mutation, sampling strategies, and also for comparatively newer studies that attempt
to join them together.  This is particularly worrisome if mutation analysis is to gain wider
acceptance among testing professionals. Further quite a few of the influential studies has been
conducted on older programming languages such as Fortran with operators specific to it, which
may not carry over to the newer and popular languages such as Java. Finally, with the modern
research into the ``do-smarter'' approach which allows byte code based mutation -- thus eliminating
the compilation step -- operators based on source code modification is no longer be applicable, and
their equivalents in the bytecode needs to be found and compared with other approaches. A detailed
discussion of these issues can be found in Section~\ref{sec:relatedwork}.

% RAHUL>AMIN: I find the non-bulleted contributions better flowing as I can work the references
% to other sections into it. Let me know if you have a better idea (As it stands neither my
% previous wording, nor your bulleted list reads good)
We have attempted to rectify this situation with a large scale study of industrial programs.
As detailed in Section~\ref{sec:methodology}, we sample 188 real world Java programs from Github, with
size ranging from 50loc to 100kloc, which makes our inferences relatively strong, and widely applicable.
We provide the results from running both original human generated test-suites, and also programmatically
generated test-suites, increasing the applicability of our findings to both researchers who seek to
compare test-suites quickly, and to industrial software testers who might be looking for a way to reduce
the time of their build cycle when using mutation coverage. The Section~\ref{sec:methodology} discusses
the sampling strategies attempted, and the operator selection strategies in detail, and the results of
our experiment are given in Section~\ref{sec:results}, followed by detailed discussion of what it implies
in section ~\ref{sec:discussion}. The research concludes in Section~\ref{sec:conclusion}.

Our specific contributions in this paper are listed as follows:

\begin{itemize}
\item Our study is the largest so far in terms of both the size of programs involved, and the number of
  programs analyzed for ``do-fewer'' strategies, which makes our inferences stronger and widely applicable.
\item We compare a large number of ``do-fewer'' strategies, especially with respect to their predictive
  power over the final mutation score.
\item Our sampling strategies include some that were suggested by other researchers, and also a few novel
  strategies that has not been published elsewhere.
\item Our recommendations are valid for both testing researchers who might be comparing programmatically
  generated test-suites, and software testing professionals working on real test-suites.
\end{itemize}
%We make the following contributions:
%\begin{itemize}
%\item We compare a large number of mutation sampling and selection strategies (Section~\ref{sec:analysis}).
%\item It seek sto improve the statistical confidence in our results by utilizing a much larger sample of projects
%      (188 projects) from the open source repository at Github~\cite{github}, using both original and automatically
%      generated test-suites using Randoop~\cite{randoop} (Section~\ref{sec:methodology}). 
%\item It analyses the results and suggest strategies (Section~\ref{sec:results}).
%\end{itemize}

%Our contributions in this paper are the following. First, as described in Section~\ref{sec:methodo%logy},
%we seek to improve the statistical confidence in our results by utilizing a much larger sample of projects
%(188 projects) from the open source repository at Github~\cite{github}, using both original and automatically
%generated test-suites using Randoop~\cite{randoop}. Second, as detailed in Section~\ref{sec:analysis}, we
%compare the reduction effectiveness of a large number of random sampling criterias (some of which are novel,
%and some that has been proposed by other researchers), and a some of the common operator selection
%criterias proposed in literature using statistical analysis. This provides us with insight for choosing
%mutation reduction techniqus according to the trade-offs in accuracy, and reduction ratios that a tester
%is willing to make for their specific requirements. This is detailed in Section~\ref{sec:results}.

%Our conclusion and recommendation for practicing testers are provided in Section~\ref{sec:conclusion}.

% \begin{comment}
% Order of most fertile operators for N-selective operator selection.
% NON_VOID_METHOD_CALLS
% NEGATE_CONDITIONALS
% REMOVE_CONDITIONALS
% DELETE_CONDITIONALS
% RETURN_VALS
% INLINE_CONSTS
% CONSTRUCTOR_CALLS
% EXPERIMENTAL_MEMBER_VARIABLE
% VOID_METHOD_CALLS
% MATH
% CONDITIONALS_BOUNDARY
% INCREMENTS
% REMOVE_INCREMENTS
% RMSWITCH
% EXPERIMENTAL_SWITCH
% INVERT_NEGS
%
% \end{comment}
%
% NON_VOID_METHOD_CALLS VOID_METHOD_CALLS CONSTRUCTOR_CALLS
% NEGATE_CONDITIONALS REMOVE_CONDITIONALS DELETE_CONDITIONALS CONDITIONALS_BOUNDARY
% RETURN_VALS INLINE_CONSTS EXPERIMENTAL_MEMBER_VARIABLE
% MATH INCREMENTS REMOVE_INCREMENTS INVERT_NEGS
% RMSWITCH EXPERIMENTAL_SWITCH