### 2014-03-13 src/related.Rnw
udpate

\section{Related Work}
\label{sec:related}
In this section we summarize the related work. There have been
different approaches to reduce the cost of mutation analysis.
These were categorized by Offutt and Untch~\cite{offutt2001mutation}
into three approaches: Do \textit{fewer}, \textit{smarter}, \textit{faster}.
The fewer approaches includes selective mutation and mutant sampling while
weak mutation, parallelization of mutation analysis, space time trade offs
are grouped under the umbrella of \textit{do smarter}. Finally
\textit{do faster} approaches include byte code based mutation analysis
systems like PIT, and similar binary patching methods.

The idea of using a subset of mutants was born along with the mutation analysis
itself. Budd~\cite{budd1980mutation} and Acree~\cite{Acree1980} showed that
even 10\% sampling can achieve 99\% accuracy for the final score.

It was further investigated by Mathur~\cite{mathur1991performance},
Wong et al.~\cite{wong1993mutation,wong1995reducing}, and
Offutt et al.~\cite{offutt1993experimental} using the Mothra~\cite{Mothra}
mutation operators for Fortran. Mathur~\cite{mathur1991performance,wong1995reducing}
suggested constrained mutation where in only two mutation operators were used which
were felt to be better than using the entire set.

% Operator Selection also assumes that fault detection ability of some operators
% are higher than others, and hence some operators should have higher weight
% than others. ~\cite{wong1997comparison}

A number of studies in the past have looked at the relative merits of operator
selection and random sampling criteria, starting with Wong and Mathur in
1995~\cite{wong1995reducing}. They compared x\% selection of each mutant type
with operator selection of just two mutation operators, and finds that both
achieved similar accuracy and reduction (80\%). This study was conducted on
five small programs in C and Fortran.

Some studies have tried to find set of \emph{sufficient mutation operators} that
while reducing the cost of mutations, maintains the correlation with the full
mutation score. Offutt~\cite{offutt1993experimental} suggested n-selective
approach which advocates step by step removal of mutation operators that produced
the largest number of mutants.

Mresa et al.\cite{mresa1999efficiency} used the cost of detection of mutants as
a means of selection, and uses it to define a set of operators. They find
that if very high mutation score close to 100\% is required, x\% selective
mutation is better than operator selection, and conversely for less stringent
scores, operator selection would be better if cost of mutant is considered.

Barbosa et. al.\cite{barbosa2001toward} provides a set of guidelines for selecting
such mutation operators. Namin et al.\cite{namin2006finding,siami2008sufficient}
formulate this problem as a variable reduction problem, and apply statistical
approaches. They found a subset of 28 operators out of 108 operators as the sufficient
operators.

Using only the statement deletion operator was first suggested by
Untch~\cite{untch2009onreduced}, who found that it correlated ($R^2 = 0.97$) with the
full mutation score better than other operator selection methods, while generating the
least number of mutants.  This was further reinforced by Deng et al.~\cite{deng2013empirical}
who defined deletion for different language elements, and found that an accuracy of 92\% is
achieved while reducing the mutants by 80\%.

Zhang et al.~\cite{zhang2010operator} compares operators based mutant selection in
~\cite{barbosa2001toward,siami2008sufficient,offutt1996anexperimental} to random mutant selection.
They find that none of these selection techniques are superior to random selection, with same number of
mutants. They also find that equal mutation sampling is more effective for larger subjects compared to
equal mutagen sampling and the reverse is true for smaller subjects.

Recently, Zhang et al.~\cite{zhang2013ase} confirmed that sampling as few as 5\% mutants
was sufficient for a very correlation (99\%) with the full mutation score, while sampling even
fewer mutants has good potential for retaining high accuracy. They investigate 8 sampling
strategies on top of operator based mutant selection and find that sampling strategies
based on program components (based on methods) performed best.

%The idea is that normal mutation has quadratic complexity on variable references while eliminatng the most
%fecund mutagens (SVR and ASR) achieves linear complexity on program size, while still approximating the full
% mutation score closely. They propose N-selective mutation which avoids N most fertile mutagens.
%Wong \cite{wong1995reducing} compares the effect of x\% mutation selection, and abs/ror constrained mutation.
%According to Wong, this is different from selective mutation in that only the most fecund operators
%are included rather than excluded. He finds that there is no evidence to suggest one over the other.
% Offutt\cite{offutt1993experimental} does not compare the effectivenss of selective mutation with x\% mutation.
% He also defines the concept of {\it operator strength} for an operator, which is defined as the total number
% of mutants that are killed by test data that is generated to kill only the mutants generated by that operator.
% He further posits that the mutagens with greatest strength may be the most useful ones. %(test data generated
% to eliminate variants of one operator may remove the mutants of other kinds too.)
%This is taken up later by Mresa\cite{mresa1999efficiency} et al. who used the cost of detection of mutants as
%a means of selection, and uses it to define another set of operators (san, aor, sdl, ror, uoi). They %also find
%that if very high mutation score close to 100\% is required, x\% selective mutation is better than operator selection,
%and conversely for less stringent scores, operator selection would be better if %const of mutant is considered.

%Offutt\cite{offutt1996anexperimental} following the previous paper\cite{offutt1993experimental} compares the
% effectiveness of selecting different operators. He divides the total operators into operand, operation, and
% statement categories, with es-selective avoiding operands, rs-selective avoiding operations, and re-selective
% avoiding statements, and e-selective only operations. He finds that using only operation based mutation operators
%  (numbering 5) is sufficient for close approximation of full mutation score. He suggests that the reason they are
%   effective is that most statements contain operators. He further posits that another reason is the amount of semantic
%change accomplished by an operator, and perhaps the most interesting ones are those that produce the smallest semantic difference.
% Replacement of Operand , Statement modification, Expression Modification: modify operators, ES-selective : not using
% operand, RS-Selective: not using expression, RE-Selective: not uinsg statement, E-selective : only expression..

%Unch\cite{untch2009onreduced} suggested using a single mutagen statement deletion. He compares the statement deletion
% with other selective mutagen sets, and finds statement deletion operator to provide the best prediction of original mutation score ($R^2 = 0.97$)
%This was carried forward by Deng et al\cite{deng2013empirical}, (according to whom, random sampling was weak when it was
%made low enough for appreciable mutant savings). They realize a mean mutation score of 92\% for adequate test suites with
% SDL alone, and generating a saving of 81\% in the number of mutants, and 41\% fewer equivalent mutants.

This work is closely related to Zhang et al.~\cite{zhang2013ase}. We extend the scope of
their study with much wider range of mutation selection and we experimented them on much more
real-world projects.

% Baseline : x\% from selected set
% MOP : x\% from each mutagen
% for these, it is sampled from operator based selected set.
% PLEm : x\% from inside the same program element.
% PLE + MOP: sample x\% from same mutagen from same element.

% Empirical Evaluation of the Statement Deletion Mutation Operator : Offutt

%\subsection{Execution Reduction}

%Weak mutation, devised by Howden \cite{howden1982weak}, considers a mutant as detected, if
%the states of program after the mutated statement in mutant and the original
%program differ. Several oher studies have refined the idea theoretically \cite{horgan1990weak}and
%observed its effectiveness emperically\cite{offutt1994empirical,Offutt:1991:SWM:120807.120826}.

%Some studies have utlized parallel computation to enhance the execution time of
%the mutation testing \cite{}. Others have used custom built compilers and interpretors \cite{}
%to reduce the overall compile time in mutation testing \cite{}.