### 2014-03-14 etc/abstract.tex
Alex edits

% 1.Statetheproblem
Mutation analysis is a well-known method for measuring the quality of
test suites. However, it is computationally intensive compared to
other measures, which makes it hard to use in practice. Choosing a
smaller subset of mutations to run is a simple approach that can
alleviate this problem.  Mutation operator selection has been heavily
researched. Recently, researchers have found that sampling mutants can
achieve accuracy and mutant reduction similar to operator selection.

%2.Saywhyit’saninterestingproblem
However, the statistical support for these conclusions has been limited, due to the small number of subject programs investigated.  The best sampling technique is also an open problem.
% that can be drawn from the previous
%research has been limited due to the small number of subject programs. Further,
%which sampling technique is best is still an open problem that needs to be
%tackled on a much larger range of test subjects.

%3.Saywhatyoursolutionachieves

Our research compares a large number of sampling and operator
selection criteria based on their ability to predict the full mutation
score as well as the consistency of mutation reduction ratios
achieved. Our results provide a ranking that can be used to choose an
appropriate mutation reduction technique by the reduction and level of fidelity to full mutation results required.

%according to the mutation reduction ratio desired, and
%how much reduction in accuracy is acceptable.

%4.Saywhatfollowsfromyoursolution

We find that all sampling approaches perform better than operator
selection methods, when considering ability to predict the full
mutation score as well as the consistency of mutation reduction ratios
achieved.

%and also the consistency of mutation reduction ratios achieved.  Specifically,
%we find that random selection of a mutant performs better than statement
%deletion operator selection criteria.

% Technical papers must be prepared in ACM conference format and must not exceed 10 pages
% (including figures and appendices but NOT including references). That is, any pages after
% the tenth must contain only references. All submissions must be in English. Submit your
% paper via the paper submission website. Submissions that do not adhere to these
% guidelines or that violate formatting will be declined without review.