### 2014-03-14 src/conclusion.Rnw
Alex changes

\section{Conclusion}
\label{sec:conclusion}

Our analysis suggests that simple random sampling --- that is, at the
project level --- works best when compared to more sophesticated schemes of
sampling. Further, we also note that random sampling performs better in
predicting final mutation score than operator selection.

This suggests that the best way to reduce the computational
requirements of mutation analysis is fortunately easy to apply in
almost any setting, given a tool that produces a good starting set of
mutants.  Namely, randomly sampling the total mutant population, even
with very small sample sizes, can effectively predict the results of
running all mutants, but at only a fraction of the computational cost.
Not only is this method simple and effective, but it is also quite
stable comapred to some alternatives, with a low standard deviation in
effectiveness across projects, for both original and random test suites.

There is a second possible point of interest here: as with coverage
metrics used to predict mutation scores~\cite{gopinath2014code}, using
a large body of actual open source code to perform experiments seems
to favor simple and easily-implemented approaches much more strongly
than limited experiments on more academic subjects.  We speculate that
the inadequacies in real world test suites, or even in randomly
generated tests for real world programs, may frequently be more simply
predicted and evaluated than for typical research subjects, which may
be smaller or involve less easily distinguished test suites.