### 2014-03-15 src/threats.Rnw
Fix space requirements

\section{Threats to Validity}
While we have taken every care to ensure that our results are valid, and to eliminate the
effects of random noise in experiments, our results are subject to the following threats
to validity.

Our results were observed on open source Java programs from the Github repository, using the
maven build system. Further, only those projects that successfully completed mutation
testing within a designated amount of time were chosen. This implies that if there is a
confounding factor in play that affects the relation between the full mutation score and
sampling mutation score, it might cause our results to be either wrong or less applicable.

Secondly, we had to rely on the PIT mutation testing tool, and had to extend its capabilities
to some extent for our purposes. Software bugs are a fact of life. While every care has been
taken to avoid them, there is still some possibility of some bugs having escaped us.