### 2014-03-16 src/results.Rnw
Shortening table description a little bit

\section{Results}
\label{sec:results}

Our results for random sampling over program elements using $sample(x)$
are given in Table~\ref{tbl:xselsample}. The result of the variation using
$round(x)$ is given in Table~\ref{tbl:xselround} and that using $ceil(x)$
is given in Table~\ref{tbl:xselceil}. The left tables show results from the
original set, while right tables show results from generated set.  The first
column in each table provides the scope of sampling: per project, per class,
per method, or per line. The second column ($\div$) provides the fraction
involved. The third column contains the $R^2$ value obtained between the
mutation score of the sampling criteria used and the full mutation score.
The fourth column ($\mu_{red}$) contains the mutant reduction factor.
This is the average of total mutants produced divided by the number of
mutants sampled by the criteria. The fifth column ($\sigma_{red}$) contains
the standard deviation of the same. A few of the projects did not return
any value for certain sampling criteria for particular mutation reduction
fractions. The sixth column contains the number of projects that actually
had valid values for the criteria being considered. The seventh column
($\beta_1$) is the average multiplier between the sampled mutation score
and the full mutation score.

Table~\ref{tbl:lcountl} provides the results for line count criteria
where the sampling was conducted after combining program elements, and
Table~\ref{tbl:lcountw} contains the results for line count criteria
where the sampling was conducted inside each program element. Similarly
Table~\ref{tbl:onesel} contains the results for the criteria of one mutant
per program element.

Finally, Table~\ref{tbl:operators} contains the results when only a single
operator is used, and Table~\ref{tbl:opsel} contains the results for specific
operator selection criteria. The first columns in these two tables contain
the particular operator or operator selection applied, and the fraction
column ($\div$) is absent. The remaining columns are same as previous tables.
The TSDL row is the result of applying only the higher order mutant equivalent
to \textit{statement deletion}.  All the \textit{N-sel} rows are the results
of N-selective operator selection strategies~\cite{offutt1993anexperimental},
using the order of operators from Figure~\ref{fig:op}. The complete data
(only for original test suites) is visualized in Figure~\ref{fig:fullimgsize}.

% \subsection{SDL vs Random selection}
% \begin{figure*}[t]
% <<sdl.x,fig.width=10, fig.height=10, out.width='.95\\linewidth', out.heigth='4in', echo=F, message=F, warn=F,fig.lp="fig:">>=
% par(mfrow=c(2,2))
% with(o.pit.all, plot(all.cov~TSDL.cov, pch='+', main='Original SDL'))
% with(o.pit.all, plot(all.cov~l.1x1_per_line.cov, pch='+', main='Original 1 per line'))
% with(r.pit.all, plot(all.cov~TSDL.cov, pch='+', main='Randoop SDl'))
% with(r.pit.all, plot(all.cov~l.1x1_per_line.cov, pch='+', main='Randoop 1 per line'))
% @
% \label{fig:sdl1}
% \caption{Statement Deletion vs One per line.}
% \end{figure*}
% <<results='asis'>>=
% oprand <- subset(stats.df,!grepl('^..op', rownames(stats.df)))
% 
% oprand.x <- subset(oprand,grepl('^.1x[0-9]+_per_', operator))
% oprand.l <- subset(oprand,!grepl('^.1x[0-9]+_per_', operator))
% @