### 2014-03-16 src/related.Rnw
Fixes from Amin

\section{Related Work}
\label{sec:related}
There are several approaches to reducing the cost of mutation analysis.
These were categorized by Offutt and Untch~\cite{offutt2001mutation}
into three approaches: do \textit{fewer}, \textit{smarter}, and \textit{faster}.
The \textit{do fewer} approaches include selective mutation and mutant sampling,
while weak mutation, parallelization of mutation analysis, and space/time trade-offs
are grouped under the umbrella of \textit{do smarter}. Finally
\textit{do faster} approaches include mutant schema generation methods,
code patching etc.

The idea of using a subset of mutants was conceived along with
mutation analysis itself. Budd~\cite{budd1980mutation} and
Acree~\cite{Acree1980} showed that even 10\% sampling can achieve 99\%
accuracy for the final score.  The idea was further investigated by
Mathur~\cite{mathur1991performance}, Wong et
al.~\cite{wong1993mutation,wong1995reducing}, and Offutt et
al.~\cite{offutt1993experimental} using the Mothra~\cite{Mothra}
mutation operators for Fortran. Mathur~\cite{mathur1991performance,wong1995reducing}
suggested constrained mutation where only two operators
were used.%, which were felt to be better than using the entire set.

A number of studies in the past have looked at the relative merits of operator
selection and random sampling criteria. Wong et al.~\cite{wong1995reducing}
compared x\% selection of each mutant type
with operator selection using just two mutation operators, and found that both
achieved similar accuracy and reduction (80\%).

Mresa et al.\cite{mresa1999efficiency} used the cost of detection of
mutants as a means of selection to define a set of operators. They
found that if very high mutation score (close to 100\%) is required,
x\% selective mutation is better than operator selection, and,
conversely, for lower scores, operator selection would be better if
the cost of mutants is considered.

Zhang et al.~\cite{zhang2010operator} compared operator based mutant
selection techniques to random mutant sampling. They found that none of
the selection techniques are superior to random sampling, with the same
number of mutants. They also found that uniform sampling of mutations is
more effective for larger subjects compared to equal sampling of mutation
operators and the reverse is true for smaller subjects.

Recently, Zhang et al.~\cite{zhang2013ase} confirmed that sampling as
few as 5\% of mutants was sufficient for a very high correlation
(99\%) with full mutation score, while sampling even fewer mutants has
good potential for retaining a high accuracy of prediction.
They investigated eight sampling strategies on top of operator-based mutant
selection and found that sampling strategies based on program components
(methods in particular) performed best.

Some studies have tried to find the set of \emph{sufficient mutation
operators} that reduce the cost of mutation but maintain correlation
with the full mutation score. Offutt et al.~\cite{offutt1993experimental}
suggested an $n$-selective approach with step-by-step removal of operators
with most numerous mutations. Barbosa et al.\cite{barbosa2001toward}
provided a set of guidelines for selecting such mutation operators.
Namin et al.\cite{namin2006finding,siami2008sufficient} formulated
the problem as a variable reduction problem, and
found that just 28 out of 108 operators in Proteum were sufficient.

Using only the statement deletion operator was first suggested by
Untch~\cite{untch2009onreduced}, who found that it had the highest
correlation ($R^2 = 0.97$) with the full mutation score compared to
other operator selection methods, while generating the smallest number of
mutants.  This was further reinforced by Deng et al.~\cite{deng2013empirical}
who defined deletion for different language elements, and found that an
accuracy of 92\% is achieved while reducing the number of mutants by 80\%.

Our work is most closely related to that of Zhang et
al.~\cite{zhang2013ase}. We extend the scope of their study with a
much wider range of mutation approaches and base our results on a much larger
set of real-world projects.