### 2014-03-17 src/results.Rnw
Update spellings, fix space etc.

\section{Results}
\label{sec:results}

Our results for random sampling over program elements using $sample(x)$
are given in Table~\ref{tbl:xselsample}. The result of the variation using
$round(x)$ is given in Table~\ref{tbl:xselround} and that using $ceil(x)$
is given in Table~\ref{tbl:xselceil}. The left tables show results from the
original set, while right tables show results from generated set.  The first
column in each table provides the scope of sampling: per project, per class,
per method, or per line. The second column ($\div$) provides the fraction
involved. The third column contains the $R^2$ value obtained between the
mutation score of the sampling criteria used and the full mutation score.
The fourth column ($\mu_{red}$) the mutant reduction factor,
which is the average of total mutants divided by the number of
mutants sampled. The fifth column ($\sigma_{red}$) is the standard deviation
of the same. As we explained in the analysis, the sixth column contains
the valid responses obtained.  The seventh column ($\beta_1$) is the
multiplier factor (coefficient reported by the regression.)

Table~\ref{tbl:lcountl} provides the results for line count criteria
where the sampling was conducted after combining program elements, and
Table~\ref{tbl:lcountw} contains the results for line count criteria
where the sampling was conducted inside each program element. Similarly
Table~\ref{tbl:onesel} contains the results for the criteria of one mutant
per program element.

Finally, Table~\ref{tbl:operators} contains the results when only a single
operator is used, and Table~\ref{tbl:opsel} contains the results for specific
operator selection criteria. The first columns in these two tables contain
the particular operator or operator selection applied, and the fraction
column ($\div$) is absent. The remaining columns are same as previous tables.
The TSDL row is the result of applying only the higher order mutant equivalent
to \textit{statement deletion}.  All the \textit{N-sel} rows are the results
of N-selective operator selection strategies~\cite{offutt1993anexperimental},
using the order of operators from Figure~\ref{fig:op}. The complete data
(only for original test suites) is visualized in Figure~\ref{fig:fullimgsize}.

% \subsection{SDL vs Random selection}
% \begin{figure*}[t]
% <<sdl.x,fig.width=10, fig.height=10, out.width='.95\\linewidth', out.heigth='4in', echo=F, message=F, warn=F,fig.lp="fig:">>=
% par(mfrow=c(2,2))
% with(o.pit.all, plot(all.cov~TSDL.cov, pch='+', main='Original SDL'))
% with(o.pit.all, plot(all.cov~l.1x1_per_line.cov, pch='+', main='Original 1 per line'))
% with(r.pit.all, plot(all.cov~TSDL.cov, pch='+', main='Randoop SDl'))
% with(r.pit.all, plot(all.cov~l.1x1_per_line.cov, pch='+', main='Randoop 1 per line'))
% @
% \label{fig:sdl1}
% \caption{Statement Deletion vs One per line.}
% \end{figure*}
% <<results='asis'>>=
% oprand <- subset(stats.df,!grepl('^..op', rownames(stats.df)))
% 
% oprand.x <- subset(oprand,grepl('^.1x[0-9]+_per_', operator))
% oprand.l <- subset(oprand,!grepl('^.1x[0-9]+_per_', operator))
% @

Our data will be available for replication on publication~\cite{data}
%\footnote{Available on publication}.