### 2015-01-12 src/paper.Rnw
update package name

% MAX LIMIT: 10 pages for all figurs and text + any number of references.
\newcommand*\per{\scalebox{.5}{\%}}
\newcommand\Mua{Mutation analysis~}
\newcommand\mua{mutation analysis~}
%\newcommand{\pitnvmc}{~NON\_\allowbreak{}VOID\_\allowbreak{}METHOD\_\allowbreak{}CALLS~}
%\newcommand{\pitnc}{~NEGATE\_\allowbreak{}CONDITIONALS~}
%\newcommand{\pitrc}{~REMOVE\_\allowbreak{}CONDITIONALS~}
%\newcommand{\pitdc}{~DELETE\_\allowbreak{}CONDITIONALS~}
%\newcommand{\pitrv}{~RETURN\_\allowbreak{}VALUES~}
%\newcommand{\pitic}{~INLINE\_\allowbreak{}CONSTANTS~}
%\newcommand{\pitcc}{~CONSTRUCTOR\_\allowbreak{}CALLS~}
%\newcommand{\pitemv}{~E\_\allowbreak{}MEMBER\_\allowbreak{}VARIABLE~}
%\newcommand{\pitvmc}{~VOID\_\allowbreak{}METHOD\_\allowbreak{}CALLS~}
%\newcommand{\pitm}{~MATH~}
%\newcommand{\pitcb}{~CONDITIONALS\_\allowbreak{}BOUNDARY~}
%\newcommand{\piti}{~INCREMENTS~}
%\newcommand{\pitri}{~REMOVE\_\allowbreak{}INCREMENTS~}
%\newcommand{\pitrs}{~RM\_\allowbreak{}SWITCH~}
%\newcommand{\pites}{~E\_\allowbreak{}SWITCH~}
%\newcommand{\pitin}{~INVERT\_\allowbreak{}NEGATIVES~}
%
\newcommand{\pitnvmc}{NMC~}
\newcommand{\pitnc}{NC~}
\newcommand{\pitrc}{RC~}
\newcommand{\pitdc}{DC~}
\newcommand{\pitrv}{RV~}
\newcommand{\pitic}{IC~}
\newcommand{\pitcc}{CC~}
\newcommand{\pitemv}{EMV~}
\newcommand{\pitvmc}{VMC~}
\newcommand{\pitm}{M~}
\newcommand{\pitcb}{CB~}
\newcommand{\piti}{I~}
\newcommand{\pitri}{RI~}
\newcommand{\pitrs}{RS~}
\newcommand{\pites}{ES~}
\newcommand{\pitin}{IN~}
% No space variants.
\newcommand{\Pitnvmc}{NMC}
\newcommand{\Pitnc}{NC}
\newcommand{\Pitrc}{RC}
\newcommand{\Pitdc}{DC}
\newcommand{\Pitrv}{RV}
\newcommand{\Pitic}{IC}
\newcommand{\Pitcc}{CC}
\newcommand{\Pitemv}{EMV}
\newcommand{\Pitvmc}{VMC}
\newcommand{\Pitm}{M}
\newcommand{\Pitcb}{CB}
\newcommand{\Piti}{I}
\newcommand{\Pitri}{RI}
\newcommand{\Pitrs}{RS}
\newcommand{\Pites}{ES}
\newcommand{\Pitin}{IN}

\begin{comment}
<<setup, include=FALSE, message=FALSE, echo=FALSE>>=
is.installed <- function(mypkg) is.element(mypkg, installed.packages()[,1])
#options(CRAN = "http://cran.stat.ucla.edu")
#if (!is.installed("xtable")) install.packages("xtable")
#if (!is.installed("ggplot2")) install.packages("ggplot2")
#if (!is.installed("reshape2")) install.packages("reshape2")

#options(warn=-1)
options(digits=3)
opts_chunk$set(fig.path='figure/twocolumn-', fig.align='center', fig.show='hold', echo=FALSE)
opts_chunk$set(cache=TRUE)
render_listings()
logit<-function(x) log(x/(1-x))
require(reshape2)
require(scales)
require(xtable)
require(ggplot2)
require(MutSel)
@
<<chunk-hook, echo=2>>=
<<setup>>
v=data(package='MutSel')
data(list = v$results[,3])
@
\end{comment}
<<>>=

f <- function(x) {
  x <- x[,c(7,8,2,3,4,5,6)]
  names(x)[names(x) == 'sampling'] <- '$\\div$'
  return(x)
}

colorize.rsqr <- function(x) {
  ifelse(x >= 0.90,
      paste("\\textcolor{blue}{", formatC(x, dig=2, format="f"), "}"),
      paste("\\textcolor{black}{", formatC(x, dig=2, format="f"), "}"))
}

colorize.it <- function(x) {
  x$sd.reduction <- ifelse(x$sd.reduction * 2 < x$mutant.reduction,
      paste("\\textcolor{black}{", formatC(x$sd.reduction, dig=2, format="f"), "}"),
      paste("\\textcolor{red}{", formatC(x$sd.reduction, dig=2, format="f"), "}"))
  x$r.square <- colorize.rsqr(x$r.square)
  return(x)
}

theme_set(theme_bw(12))
myopts <- theme(panel.background = element_blank())

@
% 10 pages for the main text + 1 for references.
<<introduction, child='introduction.Rnw'>>=
@
<<related, child='related.Rnw'>>=
@

<<methodology, child='methodology.Rnw'>>=
@

<<analysis, child='analysis.Rnw'>>=
@
<<results, child='results.Rnw'>>=
@
<<results, child='discussion.Rnw'>>=
@
<<threats, child='threats.Rnw'>>=
@
<<conclusion, child='conclusion.Rnw'>>=
@
%\FloatBarrier
%\clearpage