### 2015-01-12 src/results.Rnw
update without errors

\section{Results}
\label{sec:results}

% \subsection{SDL vs Random selection}
% \begin{figure*}[t]
% <<sdl.x,fig.width=10, fig.height=10, out.width='.95\\linewidth', out.heigth='4in', echo=F, message=F, warn=F,fig.lp="fig:">>=
% par(mfrow=c(2,2))
% with(o.pit.all, plot(all.cov~TSDL.cov, pch='+', main='Original SDL'))
% with(o.pit.all, plot(all.cov~l.1x1_per_line.cov, pch='+', main='Original 1 per line'))
% with(r.pit.all, plot(all.cov~TSDL.cov, pch='+', main='Randoop SDl'))
% with(r.pit.all, plot(all.cov~l.1x1_per_line.cov, pch='+', main='Randoop 1 per line'))
% @
% \label{fig:sdl1}
% \caption{Statement Deletion vs One per line.}
% \end{figure*}
% <<results='asis'>>=
% oprand <- subset(stats.df,!grepl('^..op', rownames(stats.df)))
% 
% oprand.x <- subset(oprand,grepl('^.1x[0-9]+_per_', operator))
% oprand.l <- subset(oprand,!grepl('^.1x[0-9]+_per_', operator))
% @

Our data will be available for replication on publication~\cite{data}
%\footnote{Available on publication}.