### 2015-01-12 src/introduction.Rnw
update

\section{Introduction}

% 1. Describe the problem % 2. State your contributions
Mutation analysis is a method for evaluating the quality of test suites. It
involves producing a family of \emph{mutants}, programs with small
differences from the original program, and evaluating the
effectiveness of test suites against these
mutants~\cite{lipton1971fault, ammann2008introduction}. Research by
Andrews et. al~\cite{andrews2005mutation} suggest that mutations can
simulate behavior of real faults.
% thus
% introduced behave in a fashion similar to real faults, with respect to
% the difficulty of detection.

An impediment to the wider adoption of mutation analysis is its
high computational cost.  The set of simple mutants for even a
moderate sized program can be very large, making mutation
analysis prohibitively time consuming.

A major strain of research into cost-reduction of mutation analysis is
to choose a smaller, representative set of
mutants~\cite{offutt2001mutation,jia2011analysis} --- often called the
\textit{do fewer} approach. This approach can be generally divided
into selective strategies and sampling strategies.

Selective mutation strategies attempt to select a representative
subset of mutation \emph{operators} based on heuristics and
statistical analysis, and apply this subset of operators to generate
mutants instead of applying the whole set of mutation
operators~\cite{offutt1993anexperimental,untch2009onreduced}.
Recent work suggests that using statement deletion alone can
be an effective approach~\cite{untch2009onreduced}.

Sampling strategies seek to randomly select a set of representative
mutants.  This was investigated first by
Acree~\cite{acree1980mutation} and Budd~\cite{budd1980mutation}, who
proposed using only $x$\% of all mutants produced. Wong and Mathur
found that random sampling with ratio as low as 10\% could provide
accurate results~\cite{jia2011analysis}.

Recent work~\cite{zhang2010icse,zhang2013ase} investigating the
relative merits of random sampling strategies and operator selection
suggests that random sampling can perform as well as or better than operator
selection, and a strategy of either sampling based on program elements or
one combining both program element-based sampling and operator selection
was the best approach.

However, as pointed out by Zhang et al.~\cite{zhang2013ase}, the field
has a serious lacuna in large scale research, both in the size of the
programs studied, and in the number and diversity of programs, which
reduces our confidence. This is true for selective mutation studies,
sampling studies, and also for comparatively newer studies that attempt
to combine methods. This is particularly worrisome if mutation analysis
is to gain wider acceptance among testing professionals.

Further, quite a few of the influential studies~\cite{offutt1993experimental,offutt1996anexperimental,wong1995reducing,mresa1999efficiency}
were conducted on older programming languages such as Fortran, with
operators specific to the language, and are not directly applicable to
newer languages such as Java.

Finally, with bytecode~\footnote{
A \textit{do faster} approach for eliminating
the compilation step to gain execution speed
} based mutation
engines like PIT~\cite{pitest} and Javalanche~\cite{javalanche}
, operators based on
source code modification are no longer applicable, and their
equivalents in bytecode need to be identified and compared with other
approaches. A detailed discussion of these issues can be found in
Section~\ref{sec:related}.

We have attempted to rectify this situation with a large scale study of real
world programs. As detailed in Section~\ref{sec:methodology}, we sample
250 Java programs from Github\footnote{Our dataset is 322 programs, but as of
this writing the larger projects have not completed finished running
full-mutation analysis.},
with size ranging from $50$ to $100,000$ lines (excluding tests), allowing
us to draw more widely applicable statistical inferences.

% For a mutation selection strategy to be considered effective, especially during
% the development, it should be able to expose as many defect-like behaviors as
% possible.
Consider any test in a test suite. It is written with the intent of
checking for the absence of some (faulty) behavior. If supposing, tests
were equivalent in the number of behaviors that they test, then the best test
suite would be one that included the largest number of non-redundant test
cases~\footnote{We show later why this assumption is valid (Specifically,
assertion distribution was $\mu=1.7$ and $\sigma=2$ )}. This suggests that
we could evaluate different mutation reduction strategies on their ability
to choose mutants such that they are killed by a larger number of
non-redundant test cases. Further, such a technique is not limited by the
adequacy of test suite under consideration.

Section~\ref{sec:methodology} discusses the sampling and operator
selection strategies we study in detail, and the results of our
experiment are given in Section~\ref{sec:results}, followed by
detailed discussion of what these results imply in Section
~\ref{sec:discussion}. The specific contributions of this paper are:

\begin{itemize}

\item Our study is the largest so far in terms of
both the size of programs involved (50 LOC to 100 KLOC excluding tests),
and the number of programs analyzed (250 unique open source projects) for
mutant reduction strategies. This allows for stronger and more widely
applicable conclusions about effectiveness.

\item We compare a much larger
number of mutant reduction strategies than previous studies. Our reduction
strategies include all the common and influential strategies for
random sampling, element scoped sampling, and operator selection, that
have been proposed in previous studeies.

\item Our evaluation crtieria is applicable to non-adequate test suites
which makes the results immediately relevant for real world programs with
non-adequate test suites.

\item Most importantly, we find that current operator reduction strategies
seldom perform better than random sampling of mutants, which suggests a need
for further research in this area.

\end{itemize}