### 2015-01-15 src/results.Rnw
results

\section{Results}
\label{sec:results}

The results for operator selection criteria are given in Table~\ref{tbl:selective}. Those of N-Selection are given in Table~\ref{tbl:nselection}. The fractional sampling per operator is provided in Table~\ref{tbl:opsample}, and that of element scoped sampling is provided in Table~\ref{tbl:element}. Finally, operation selection using just single operators are given in Table~\ref{tbl:singles}.

Our data is available for replication~\cite{ourdata}.