### 2015-01-15 src/discussion.Rnw
nuke discussion

\section{Discussion}
\label{sec:discussion}

To be added.