### 2015-01-15 src/threats.Rnw
update

\section{Threats to Validity}
While we have taken every care to ensure that our results are valid, and
to eliminate the effects of random noise, our results are subject to the
following threats to validity.

Our results were observed on open source Java programs from the Github
repository, and Apache Commons Library, using the maven build system. Further,
only those projects that successfully completed mutation testing within
a designated amount of time were chosen.  This implies that if there is
a confounding factor in play that affects the relation between the full
mutation score and sampling mutation score, it might cause our results to
be either wrong or less applicable.

Secondly, we had to rely on the PIT mutation testing tool, and had to extend
its capabilities to some extent for our purposes. Software bugs are a fact
of life. While every care has been taken to avoid them, there is still some
possibility of some bugs having escaped us.

Finally, and most importanly, most of the previous research has been done on
source level mutants. However, for a study of the size we have conducted,
it is infeasible to rely on source level mutants (even using the bytecode
based fast mutation evaluation provided by PIT, the mutation analysis took
months to complete). This also means that we had to rely on translation of
source level mutants to their bytecode equivalents, which was in most cases
not a one to one process, and this could have introduced unintented errors.
However, as we mentioned earlier, it is infeasible to run mutation analysis
on source level for large programs using the tools we currently have. It is
also important to evaluate different strategies on byte code based engines
at least to provide guidance for testers trying to do the same.