### 2015-01-17 src/fig.mutantdist.Rnw

<<mutantdet,fig.width=10, fig.height=5, out.width='.95\\linewidth', out.heigth='4in', echo=F, message=F, warn=F,fig.lp="fig:">>=
bcol <- '#2171B5'
icol <- '#9ECAE0'
m.db <- subset(allmutants.db, total > 0)
d.db <- subset(allmutants.db, detected > 0)
m.db$mutations <- m.db$mtotal * 100.0 / m.db$total
d.db$detected <- d.db$mdetected * 100.0 / d.db$detected

tmp1 <- subset(m.db, select=c("mutator","mutations"))
tmp1$type <- "mutations"
tmp2 <- subset(d.db, select=c("mutator","detected"))
colnames(tmp2) <- c("mutator","mutations")
tmp2$type <- "detected"
tmp <- rbind(tmp1, tmp2)
tmp$type <- factor(tmp$type, levels=c("mutations", "detected"))
tmp$mutator <- factor(tmp$mutator, levels=c("nmc", "rv", "ic", "dc", "nc", "rc", "vmc", "cc", "emv", "m", "cb", "i", "ri", "rs", "es", "in"))
options(warn=-1)
ggplot(aes(y=mutations, x=mutator, fill=type),data=tmp) +
  geom_boxplot(outlier.color=NA, outlier.shape=NA) +
  scale_fill_manual(name = "Mutation Details", values = c(bcol, icol)
                    , labels = c("mutations" = "Total mutations", "detected" = "Detected mutations")) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
 theme(axis.line = element_line(colour = "black"))
 # coord_flip() +
options(warn=0)
# + stat_boxplot(geom ='errorbar') 
#p1 <- ggplot(aes(y=mutations, x=mutator, fill=F),data=m.db) + geom_boxplot() + guides(fill=FALSE) + coord_flip()
#p2 <- ggplot(aes(y=detected, x=mutator, fill=F),data=d.db) + geom_boxplot() + guides(fill=FALSE) + coord_flip()

#grid.newpage()
#pushViewport(viewport(layout = grid.layout(1, 2)))
#print(p1, vp = vplayout(1, 1))
#print(p2, vp = vplayout(1, 2))
@