### 2015-01-17 src/fig.testdist.Rnw

<<testdet,fig.width=10, fig.height=5, out.width='.95\\linewidth', out.heigth='4in', echo=F, message=F, warn=F,fig.lp="fig:">>=
bcol <- '#2171B5'
icol <- '#9ECAE0'
m.db <- subset(allmutants.db, ncovering > 0)
d.db <- subset(allmutants.db, nkilling > 0)
#m.db$covering <- m.db$tstcovering * 100.0 / m.db$ncovering
m.db$covering <- m.db$tstcovering * 100.0 / m.db$ntests
#d.db$killing <- d.db$tstkilling * 100.0 / d.db$nkilling
d.db$killing <- d.db$tstkilling * 100.0 / d.db$ntests

tmp1 <- subset(m.db, select=c("mutator","covering"))
colnames(tmp1) <- c("mutator","tests")
tmp1$type <- "covering"

tmp2 <- subset(d.db, select=c("mutator","killing"))
colnames(tmp2) <- c("mutator","tests")
tmp2$type <- "killing"

tmp <- rbind(tmp1, tmp2)
tmp$type <- factor(tmp$type, levels=c("covering", "killing"))
tmp$mutator <- factor(tmp$mutator, levels=c("nmc", "rv", "ic", "dc", "nc", "rc", "vmc", "cc", "emv", "m", "cb", "i", "ri", "rs", "es", "in"))
options(warn=-1)
ggplot(aes(y=tests, x=mutator, fill=type),data=tmp) +
  geom_boxplot(outlier.color=NA, outlier.shape=NA) +
  scale_fill_manual(name = "Mutation Details", values = c(bcol, icol)
                    , labels = c("covering" = "Total covering", "killing" = "Total killing")) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
 theme(axis.line = element_line(colour = "black"))
options(warn=0)
@