### 2015-01-17 src/tables.Rnw

\begin{table*}[!ht]
%\caption*{Selective}
<<results='asis'>>=
eff <- function(v) {
  v$strategy.effectiveness <- v$strategy.effectiveness - v$baseline.effectiveness
  return(v)
}
tbl.place <- function(xtb) {
  align(xtb) <- "p{0.1\\textwidth}|p{0.15\\textwidth}|p{0.15\\textwidth}|p{0.15\\textwidth}|p{0.15\\textwidth}|p{0.15\\textwidth}"
  return(xtb)
}
print(tbl.place(xtable(eff(s.selective.db), caption='Selective Strategy', label='table:selective')),include.rownames=T, floating=F, size='\\small',sanitize.text.function=function(x){x})
@
\caption {The operator selection strategy}
\label{tbl:selective}
\end{table*}

\begin{table*}[!ht]
%\caption*{N-Selection}
<<results='asis'>>=
print(tbl.place(xtable(eff(s.nselection.db), caption='N-Selection', label='table:nselection')), include.rownames=T, floating=F, size='\\small',sanitize.text.function=function(x){x})
@
\caption {The N-selective strategy}
\label{tbl:nselection}
\end{table*}

\begin{table*}[!ht]
%\caption*{Operator Samples x\%}
<<results='asis'>>=
print(tbl.place(xtable(eff(s.opsample.db), caption='Operator x samples', label='table:opsample')), include.rownames=T, floating=F, size='\\small',sanitize.text.function=function(x){x})
@
\caption {The operator based x\% sample strategy}
\label{tbl:opsample}
\end{table*}

\begin{table*}[!ht]
%\caption*{Element Scoped}
<<results='asis'>>=
print(tbl.place(xtable(eff(s.lmcp.db), caption='Element Scoped', label='table:element')), include.rownames=T, floating=F, size='\\small',sanitize.text.function=function(x){x})
@
\caption {The element based x\% sample strategy}
\label{tbl:element}
\end{table*}

\begin{table*}[!ht]
<<tblsingles, child='tbl.singleop.Rnw'>>=
@
%\caption*{Single Operators}
\caption {The single operator selection strategy}
\label{tbl:singles}
\end{table*}