### 2015-01-17 src/ttests.Rnw
update

<<tests,results='asis'>>=
mintc.t.test <- function(db) {
  return(with(db, t.test(sel.mintc,sample.mintc, paired=T))) #, alternative=c("greater"))))
}
mutant.t.test <- function(db) {
  return(with(db, t.test(sel.all.killtcd_mutant,sample.all.killtcd_mutant, paired=T))) #, alternative=c("greater"))))
}
db <- subset(common.db, kind == "nsel")
nsel.t <- mintc.t.test(db)
nsel.m <- mutant.t.test(db)
#-- Insufficient evidence that sel.mintc is greater than random selection
db <- subset(common.db, kind == "opsel")
opsel.t <- mintc.t.test(db)
opsel.m <- mutant.t.test(db)
#-- Insufficient evidence that sel.mintc is greater than random selection
db <- subset(common.db, kind == "single")
single.t <- mintc.t.test(db)
single.m <- mutant.t.test(db)
#-- Insufficient evidence that sel.mintc is greater than random selection
db <- subset(common.db, kind == "opx")
opx.t <- mintc.t.test(db)
opx.m <- mutant.t.test(db)
#-- Insufficient evidence that sel.mintc is greater than random selection
db <- subset(common.db, kind == "lmcp")
lmcp.t <- mintc.t.test(db)
lmcp.m <- mutant.t.test(db)
# #---------------------------------------
# # nmc     nc       rc       dc       rv       ic       cc
# # emv      vmc      m       cb       i        ri       rs       es in
# db <- subset(common.db, operator == "rm_nmc")
# with(db, t.test(sel.mintc,sample.mintc, paired=T, alternative=c("greater")))
# with(db, t.test(sel.all.killtcd_mutant,sample.all.killtcd_mutant, paired=T, alternative=c("greater")))
# #---------------------------------------
# db <- subset(common.db, operator == "cc")
# with(db, t.test(sel.mintc,sample.mintc, paired=T, alternative=c("greater")))
# with(db, t.test(sel.all.killtcd_mutant,sample.all.killtcd_mutant, paired=T, alternative=c("greater")))
# #---------------------------------------
# db <- subset(common.db, operator == "clazz")
# with(db, t.test(sel.mintc,sample.mintc, paired=T)) #, alternative=c("greater")))
# with(db, t.test(sel.all.killtcd_mutant,sample.all.killtcd_mutant, paired=T))#, alternative=c("greater")))
db <- subset(common.db, operator=="constrained")
constrained.t <- mintc.t.test(db)
constrained.m <- mutant.t.test(db)

db <- subset(common.db, operator=="eselective")
eselective.t <- mintc.t.test(db)
eselective.m <- mutant.t.test(db)

db <- subset(common.db, operator=="javalanche")
javalanche.t <- mintc.t.test(db)
javalanche.m <- mutant.t.test(db)

db <- subset(common.db, operator=="varreduction")
variable.t <- mintc.t.test(db)
variable.m <- mutant.t.test(db)

db <- subset(common.db, operator=="sdl")
sdl.t <- mintc.t.test(db)
sdl.m <- mutant.t.test(db)

db <- subset(common.db, operator=="line")
line.t <- mintc.t.test(db)
line.m <- mutant.t.test(db)

db <- subset(common.db, operator=="method")
method.t <- mintc.t.test(db)
method.m <- mutant.t.test(db)

db <- subset(common.db, operator=="clazz")
clazz.t <- mintc.t.test(db)
clazz.m <- mutant.t.test(db)

# rm_none      rm_nmc       rm_nc        rm_rc        rm_dc        rm_rv
# rm_ic        rm_cc       rm_emv       rm_vmc       rm_m         rm_cb
# rm_i         rm_ri        rm_rs        rm_es       
db <- subset(common.db, operator=="rm_nmc")
rm.nmc.t <- mintc.t.test(db)
rm.nmc.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_rv")
rm.rv.t <- mintc.t.test(db)
rm.rv.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_ic")
rm.ic.t <- mintc.t.test(db)
rm.ic.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_dc")
rm.dc.t <- mintc.t.test(db)
rm.dc.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_nc")
rm.nc.t <- mintc.t.test(db)
rm.nc.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_rc")
rm.rc.t <- mintc.t.test(db)
rm.rc.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_vmc")
rm.vmc.t <- mintc.t.test(db)
rm.vmc.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_cc")
rm.cc.t <- mintc.t.test(db)
rm.cc.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_emv")
rm.emv.t <- mintc.t.test(db)
rm.emv.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_m")
rm.m.t <- mintc.t.test(db)
rm.m.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_cb")
rm.cb.t <- mintc.t.test(db)
rm.cb.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_i")
rm.i.t <- mintc.t.test(db)
rm.i.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_ri")
rm.ri.t <- mintc.t.test(db)
rm.ri.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_rs")
rm.rs.t <- mintc.t.test(db)
rm.rs.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_es")
rm.es.t <- mintc.t.test(db)
rm.es.m <- mutant.t.test(db)

db <- subset(common.db, operator=="rm_in")
rm.in.t <- mintc.t.test(db)
rm.in.m <- mutant.t.test(db)

@