### 2015-01-19 src/r.setup.Rnw

\begin{comment}
<<setup, include=FALSE, message=FALSE, echo=FALSE>>=
is.installed <- function(mypkg) is.element(mypkg, installed.packages()[,1])
#options(CRAN = "http://cran.stat.ucla.edu")
#if (!is.installed("xtable")) install.packages("xtable")
#if (!is.installed("ggplot2")) install.packages("ggplot2")
#if (!is.installed("reshape2")) install.packages("reshape2")

#options(warn=-1)
options(digits=3)
opts_chunk$set(fig.path='figure/twocolumn-', fig.align='center', fig.show='hold', echo=FALSE)
opts_chunk$set(cache=TRUE)
render_listings()
logit<-function(x) log(x/(1-x))
require(reshape2)
require(scales)
require(xtable)
require(ggplot2)
require(grid)
require(MutSel)
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
theme = theme_set(theme_minimal())
@
<<chunk-hook, echo=2>>=
<<setup>>
v=data(package='MutSel')
data(list = v$results[,3])
@
\end{comment}
<<>>=

f <- function(x) {
  x <- x[,c(7,8,2,3,4,5,6)]
  names(x)[names(x) == 'sampling'] <- '$\\div$'
  return(x)
}

colorize.rsqr <- function(x) {
  ifelse(x >= 0.90,
      paste("\\textcolor{blue}{", formatC(x, dig=2, format="f"), "}"),
      paste("\\textcolor{black}{", formatC(x, dig=2, format="f"), "}"))
}

colorize.it <- function(x) {
  x$sd.reduction <- ifelse(x$sd.reduction * 2 < x$mutant.reduction,
      paste("\\textcolor{black}{", formatC(x$sd.reduction, dig=2, format="f"), "}"),
      paste("\\textcolor{red}{", formatC(x$sd.reduction, dig=2, format="f"), "}"))
  x$r.square <- colorize.rsqr(x$r.square)
  return(x)
}

theme_set(theme_bw(12))
myopts <- theme(panel.background = element_blank())

@