### 2015-01-19 src/tbl.pitmut.Rnw
updat

\begin{table}
\small{
\begin{tabular}{ll}
\hline \hline \\
\Pitin & Remove negative sign from numbers\\
\Pitrv & Mutate return values\\
\Pitm & Mutate arithmetic operators\\
\Pitvmc & Remove void method calls\\
\Pitnc & Negate conditional statements\\
\Pitcb & Modify boundaries in logical conditions\\
\Piti & Modify increment and decrement statements\\
\Pitnvmc & Remove non-void method calls, returning default value \\
\Pitcc & Replace constructor calls, returning null \\
\Pitic & Replace inline constants with default value \\
\Pitri & Remove increment and decrement statements \\
\Pitemv & Replace member variable assignments with default value \\
\Pites & Modify switch statements \\
\Pitrs & Replace switch labels with default (thus removing them) \\
\Pitrc & Replace boolean conditions with true\\
\Pitnc & Replace boolean conditions with false\\
\hline
\end{tabular}
}
\caption{PIT Mutation Operators (We have used abbreviations instead of operator names.)}
\label{tbl:pitop}
\end{table}