### 2015-01-20 src/conclusion.Rnw
conclusion

\section{Conclusion}
\label{sec:conclusion}

Our analysis suggests that while some of the different strategies produce
slightly advantageous mutation subsets, overall, the advantage thus gained
is very little. None of the common operator selection strategies performed
consistently well against random sampling while using either the utility or
the mutation effectiveness criteria. While some of the individual operators
seemed to indicate a mean positive test utility or a mean strategy
effectiveness, the effect itself is too small (of the order of one or two
mutants or tests) to be of much help for a tester in the real world.

A similar result seems to be the case for operator based x\% selection
strategy or even for the element based x\% strategy. They in general show a
slight advantage in test utility which while consistent, is not of much
practical utility.

This suggests that that none of the well known reduction strategies are
actually able to consistently select useful mutants, and the best path forward
for a working tester is to rely on the simplest scheme of all -- the random
selection of mutants.

There is a second possible point of interest here: as with coverage
metrics used to predict mutation scores~\cite{gopinath2014code}, using
a large body of actual open source code to perform experiments seems
to favor simple and easily-implemented approaches much more strongly
than limited experiments on more academic subjects.  We speculate that
the inadequacies in real world test suites may frequently be more simply
predicted and evaluated than for typical research subjects, which may
be smaller or involve less easily distinguished test suites.
% TODO: Should we add this.
%\begin{comment}

There is also an inherent danger to using only a small sample of same academic
subjects over and over again. We may tend to see patterns that are specific
to the subjects, which are not present in the wider population. The
non-performance of the popular strategies against a wider dataset such as
ours may also be seen as a possible symptom of this over-fitting.
%\end{comment}